Characterization and Photometric Performance of the Hyper Suprime-Cam Software Pipeline.

### 2017-04-24 draft/ms_v4.tex
More minor adjustments

    In February 2017, the HSC collaboration released the first 1.7 years of data to
    the public (DR1;
    \citealt{HSCDR1})\footnote{\texttt{https://hsc-release.mtk.nao.ac.jp }}.
    For our work, we focused on data from the Wide layer as they relate to key
    scientific goals of the HSC survey, which include weak lensing cosmology, galaxy evolution, and studies of galaxy clusters.  The Wide layer data in DR1 corresponds to ${\sim}108$ deg$^2$ spread over 
   six fields (XMM--LSS, GAMA09H, WIDE12H, GAMA15H, HECTOMAP, and VVDS)
    \footnote{A small AEGIS field is also observed for calibration purposes.}.
    Except for the HECTOMAP field, the regions are all close to the equator.
    In the Wide layer, the $g$ and $r$ bands have 10 minute exposures broken into four
    dithers.  The  $i$, $z$, and $y$ bands have 20 minute exposures broken into six dithers. The survey prioritizes the observations so that the $i$-band has the best seeing conditions to improve galaxy shape measurements for weak lensing science. The  $i$ band data in the Wide layer reach a $5\sigma$ point source limiting magnitude of  $i{\sim} 26.4$ mag  and have a median seeing with FWHM${\sim}0.6$\asec{}. \citealt{HSCDR1} provides details on the first data release and data status.

@@

The HSC data are reduced using a pipeline that builds on the prototype pipeline being designed by the Large Synoptic Survey Telescope’s Data Management system (\citealt{Ivezic2008, Axelrod2010,
    Juric2015}) and is described in Bosch\etal (in prep.). DR1 data are reduced by \hscpipe{} \texttt{v4.0.5}. Because\synpipe{} is intrinsically connected to the complex data reduction
    processes in \hscpipe{}, we briefly introduce the main steps below, and we introduce a few key HSC/LSST terms.

@@

        \item \textbf{HSC Photometry}:
            The multi-band catalogs generated by \hscpipe{} contain various types of photometry measurements (see \citealt{HSCDR1} for details). Here, we focus on characterizing the the \forced{} \texttt{psf} and \cmodel{}
            photometry. The
            \texttt{psf} photometry should provide the most appropriate magnitudes and
            colors for point sources. The  \cmodel{} should be used for galaxies. The HSC  \cmodel{} algorithm is described in Bosch\etal, in prep. and is an  improved version of the SDSS \cmodel{} magnitude (\citealt{Lupton2001, Abazajian2004}). The HSC  \cmodel{} fits objects using both a de Vaucouleurs and an exponential profiles taking into account the smearing by the PSF.  \cmodel{} then linearly combines the two results to find the best fit to the galaxy image. \cmodel{} magnitudes are designed to yield accurate fluxes and colors for galaxies. Star--galaxy separation is performed based on
            the difference between \texttt{psf} and \cmodel{}{} magnitudes.

@@

        \item \textbf{Authentic Data Processing}:
            To follow the data reduction process as realistically as possible, we choose to
            start from the single-\visit{} images instead of directly putting synthetic
            objects on the final \coadd{} images (e.g., \citealt{Suchyta2016}).
            Through this approach, subtle effects like seeing differences among
            \visits{} and small errors in astrometric calibration can be taken
            into account.
            Every synthetic object will experience all the steps we would use for a real object: detection, deblending, stacking, and measurement.
            This can be important for challenging tasks like WL measurements
            or the detection of high-$z$ galaxies.
            The downside of this choice is that it slows down the overall run-time because we need to inject synthetic objects into all the \visits{} that contribute to a \tract{}.
        \item \textbf{Flexible Capabilities}: \synpipe{} is designed to be flexible enough to be useful for HSC users with a range of scientific goals.
            We provide default catalogs to generate samples of synthetic stars
            and galaxies with realistic magnitudes and color distributions along with
            tools to help the user work on HSC DR1 data. However, \synpipe{} also easily allows users to  design and use their own custom made input catalogs.
            \synpipe{} is already being used in a wide range of
            scientific applications (see Section \ref{sec:summary} for details).

@@

        \item \textbf{Input catalog}. This is a catalog in \texttt{FITS} (Flexible Image Transport System)
            format that contains the  coordinates, magnitudes, and model parameters of synthetic  galaxies.  The positions of synthetic galaxies can be specified  via the input catalog, they can be distributed randomly over a given region, or on a grid (the grid option is useful in that it avoids blends between Synthetic objects, Murata\etal in prep.).  
    \end{itemize}

Synthetic objects can be one of the following:

@@

        \item \textbf{Point source}: \synpipe{} simply uses the PSF model from \hscpipe{} as the model for
            a point source (stars and/or quasi-stellar objects).
            \hscpipe{} uses a special version of \texttt{PSFEx}
            (\citealt{Bertin2011, Bertin2013}) to characterize PSF as a function
            of position and the PSF model can be reconstructed
            for any location on the image. To inject point sources, the user only needs to specify a catalog of coordinates and magnitudes.
        \item \textbf{\ser{}models}:
            \texttt{SynPipe} uses \galsim{} \texttt{v1.4} (\citealt{Rowe2015}) to
            simulate galaxies.
            \galsim{} is an image simulation tool that was designed for the GRavitational lEnsing Accuracy Testing 3 (GREAT3) challenge
            (\citealt{Mandelbaum2014}). Currently, \synpipe{} allows synthetic galaxies to be modeled by a single
            or a double \citet{Sersic1963}
            profile\footnote{More flexible model choices will be added later.}. The \ser{} profile is flexible enough to describe the overall flux
            distributions of galaxies near and far, both early-type or late-type.
            A double-\ser{} model can simulate a galaxy with even more realistic
            structural details (e.g., bulge$+$disk).
            For each \ser{} component, the user needs to provide the magnitude,
            effective radius ($R_{\mathrm{e}}$, in unit of arcsec), \ser{} index
            ($n_{\mathrm{Ser}}$), axis ratio ($\mathrm{b}/\mathrm{a}$), and position
            angle (PA).
            External shear can also be applied to a synthetic galaxy ($g_1$ and
            $g_2$, see \citealt{Rowe2015} for more details).
        \item \textbf{Galaxies from the GREAT3 challenge}: \synpipe{} also allows users to choose from parametric models or real
            high-resolution galaxy images used in the GREAT3 WL challenge. Real galaxy images are drawn from the \hst{}/ACS F814W ($I$-band) images of galaxies
            in the COSMOS field (e.g., \citealt{Scoville2007,Leauthaud2007}). Instead of using real galaxy images, the users may also use models of  COSMOS galaxies to $I_{\mathrm{Auto}}\leq25.2$ mag via a method developed by
            \citet{Lackner2012}.  Parametric models of COSMOS galaxies include a single-\ser{} and the \ser{}
            bulge$+$Exponential disk model. The parameter values for these models are stored in the
            \texttt{GalSim.COSMOSCatalog()} catalog.
            To access them, the user can provide the ID number of COSMOS galaxies
            in the input catalog, or can simply opt to let \synpipe{} randomly select galaxies from the COSMOS input catalog.
    \end{itemize}
%% ------------------------------------------------------------------------------------ %%

@@

        If necessary, an additional shear ($g_1$ and $g_2$) can be applied to the 
        model at this point.
        Then, \synpipe{} passes the reconstructed PSF image into a
        \galsim{}\texttt{InterpolatedImage} object and convolves it with the galaxy 
        model using \galsim{}\texttt{.Convolve}.
        After this, \synpipe{} uses the \galsim{}\texttt{.drawImage} method to create 
        the image of the simulated galaxy.
        A component with a high \ser{} index ($n_{\mathrm{Ser}} > 4$) often requires 
        a large image size to cover all of its flux, so \synpipe{} allows the user 
        to truncate the model at a given radius (N$\times R_{\mathrm{e}}$).
        \synpipe{} will also give warning information when a component with $
        n_{\mathrm{Ser}} > 6$ is encountered because it takes much longer to achieve
        accurate PSF convolution in such a model; and the result is, therefore,
        not a realistic galaxy model.

@@

    \todo{Details for stars and galaxies}

>>>>

    In February 2017, the HSC collaboration released the first 1.7 years of data to
    the public (DR1;
    \citealt{HSCDR1})\footnote{\texttt{https://hsc-release.mtk.nao.ac.jp }}.
    For our work, we focused on data from the Wide layer as they relate to key
    scientific goals of the HSC survey, which include weak lensing cosmology, galaxy 
    evolution, and studies of galaxy clusters. 
    The Wide layer data in DR1 correspond to ${\sim}108$ deg$^2$ spread over 
    six fields (XMM--LSS, GAMA09H, WIDE12H, GAMA15H, HECTOMAP, and VVDS)
    \footnote{A small AEGIS field is also observed for calibration purposes.}.
    Except for the HECTOMAP field, the regions are all close to the equator.
    In the Wide layer, the $g$ and $r$ bands have 10 minute exposures broken into 
    four dithers.  
    The $i$, $z$, and $y$ bands have 20 minute exposures broken into six dithers. 
    The survey prioritizes the observations so that the $i$-band has the best seeing 
    conditions to improve galaxy shape measurements for weak lensing science. 
    The $i$ band data in the Wide layer reach a $5\sigma$ point source limiting 
    magnitude of $i{\sim} 26.4$ mag and have a median seeing with 
    FWHM${\sim}0.6$\asec{}. 
    \citealt{HSCDR1} provides details on the first data release and data status.

@@

    The HSC data are reduced using a pipeline that builds on the prototype pipeline 
    being designed for the Large Synoptic Survey Telescope’s Data Management system
    (\citealt{Ivezic2008, Axelrod2010, Juric2015}) and is described in 
    Bosch\etal (in prep.). 
    DR1 data are reduced by \hscpipe{} \texttt{v4.0.5}. 
    Because\synpipe{} is intrinsically connected to the complex data reduction
    processes in \hscpipe{}, we briefly introduce the main steps below along with a 
    few key HSC/LSST terms.

@@

        \item \textbf{HSC Photometry}:
            The multi-band catalogs generated by \hscpipe{} contain various types of 
            photometric measurements (see \citealt{HSCDR1} for details). 
            Here, we focus on characterizing the the \forced{} \texttt{psf} and 
            \cmodel{} photometry. 
            The \texttt{psf} photometry should provide the most appropriate magnitudes 
            and colors for point sources, while the \cmodel{} should be used for 
            galaxies. 
            The HSC \cmodel{} algorithm is described in Bosch\etal, in prep. and is an
            improved version of the SDSS \cmodel{} magnitude (\citealt{Lupton2001, 
            Abazajian2004}). 
            The HSC \cmodel{} fits objects using both a de~Vaucouleurs and an 
            exponential profiles taking into account the smearing by the PSF.  
            \cmodel{} then linearly combines the two results to find the best fit to 
            the galaxy image. 
            \cmodel{} magnitudes are designed to yield accurate fluxes and colors for 
            galaxies. 
            Star--galaxy separation is performed based on the difference between 
            \texttt{psf} and \cmodel{}{} magnitudes.

@@

        \item \textbf{Authentic Data Processing}:
            To follow the data reduction process as realistically as possible, we 
            choose to start from the single-\visit{} images instead of directly putting 
            synthetic objects on the final \coadd{} images 
            (e.g., \citealt{Suchyta2016}).
            Through this approach, subtle effects like seeing differences among
            \visits{} and small errors in astrometric calibration can be taken
            into account.
            Every synthetic object will experience all the steps we would use for a 
            real object: detection, deblending, stacking, and measurement.
            This can be important for challenging tasks like WL measurements or the 
            detection of high-$z$ galaxies.
            The downside of this choice is that it slows down the overall run-time
            because we need to inject synthetic objects into all the \visits{} that 
            contribute to a \tract{}.

        \item \textbf{Flexible Capabilities}: 
            \synpipe{} is designed to be flexible enough to be useful for HSC users 
            with a range of scientific goals.
            We provide default catalogs to generate samples of synthetic stars
            and galaxies with realistic magnitudes and color distributions along with
            tools to help the user work on HSC DR1 data. 
            However, \synpipe{} also easily allows users to design and use their own 
            custom made input catalogs.
            \synpipe{} is already being used in a wide range of
            scientific applications (see Section \ref{sec:summary} for details).

@@

        \item \textbf{Input catalog}. 
            This is a catalog in \texttt{FITS} (Flexible Image Transport System)
            format that contains the coordinates, magnitudes, and model parameters of 
            synthetic objects.  
            The positions of synthetic galaxies can be specified via the input catalog, 
            they can be distributed randomly over a given region, or on a grid 
            (the grid option is useful in that it avoids blends between Synthetic 
            objects, e.g. see Murata\etal in prep.).

    \noindent Synthetic objects can be one of the following:

@@

        \item \textbf{Point source}: 
            \synpipe{} simply uses the PSF model from \hscpipe{} as the model for
            a point source (stars and/or quasi-stellar objects).
            \hscpipe{} uses a special version of \texttt{PSFEx}
            (\citealt{Bertin2011, Bertin2013}) to characterize PSF as a function
            of position and the PSF model can be reconstructed for any location on 
            the image. 
            To inject point sources, the user only needs to specify a catalog of 
            coordinates and magnitudes.

        \item \textbf{\ser{}models}:
            \texttt{SynPipe} uses \galsim{} \texttt{v1.4} (\citealt{Rowe2015}) to
            simulate galaxies.
            \galsim{} is an image simulation tool that was designed for the 
            GRavitational lEnsing Accuracy Testing 3 (GREAT3) challenge
            (\citealt{Mandelbaum2014}). 
            Currently, \synpipe{} allows synthetic galaxies to be modeled by a
            single or a double \citet{Sersic1963}
            profile\footnote{More flexible model choices will be added later.}. 
            The \ser{} profile is flexible enough to describe the overall flux
            distributions of galaxies near and far, both early-type or late-type.
            A double-\ser{} model can simulate a galaxy with even more realistic
            structural details (e.g., bulge$+$disk).
            For each \ser{} component, the user needs to provide the magnitude,
            effective radius ($R_{\mathrm{e}}$, in unit of arcsec), \ser{} index
            ($n_{\mathrm{Ser}}$), axis ratio ($\mathrm{b}/\mathrm{a}$), and position
            angle (PA).
            External shear can also be applied to a synthetic galaxy ($g_1$ and
            $g_2$, see \citealt{Rowe2015} for more details).

        \item \textbf{Galaxies from the GREAT3 challenge}: 
            \synpipe{} also allows users to choose from parametric models or real
            high-resolution galaxy images used in the GREAT3 WL challenge. 
            Real galaxy images are drawn from the \hst{}/ACS F814W ($I$-band) images 
            of galaxies in the COSMOS field (e.g., \citealt{Scoville2007,Leauthaud2007}). 
            Instead of using real galaxy images, the users may also use models of COSMOS 
            galaxies to $I_{\mathrm{Auto}}\leq25.2$ mag via a method developed by
            \citet{Lackner2012}.  
            Parametric models of COSMOS galaxies include a single-\ser{} and the \ser{}
            bulge$+$Exponential disk model. 
            The parameter values for these models are stored in the
            \texttt{GalSim.COSMOSCatalog()} catalog.
            To access them, the user can provide the ID number of COSMOS galaxies
            in the input catalog, or can simply opt to let \synpipe{} randomly select 
            galaxies from the COSMOS input catalog.

@@

%% ------------------------------------------------------------------------------------ %%            
        If necessary, an additional shear ($g_1$ and $g_2$) can be applied to the 
        model at this point.
        Then, \synpipe{} passes the reconstructed PSF image into a
        \galsim{}\texttt{InterpolatedImage} object and convolves it with the galaxy 
        model using \galsim{}\texttt{.Convolve}.
        After this, \synpipe{} uses the \galsim{}\texttt{.drawImage} method to create 
        the image of the simulated galaxy.
        A component with a high \ser{} index ($n_{\mathrm{Ser}} > 4$) often requires 
        a large image size to cover all of its flux, so \synpipe{} allows the user 
        to truncate the model at a given radius (N$\times R_{\mathrm{e}}$).
        \synpipe{} will also give warning information when a component with $
        n_{\mathrm{Ser}} > 6$ is encountered because it takes much longer to achieve
        accurate PSF convolution in such a model; and the result is, therefore,
        not a realistic galaxy model.

@@
